---
title: "Calculating covariance"
author: "Lasse Hansen"
date: "10/9/2019"
output: html_document
---

Making the data frame 

```{r}
library(ggplot2)
Sarah <- c(1.95,1.58,1.70,2.46,2.27,2.62,3.32,3.51,3.89,3.41)
Mother <- c(3.21,4.04,3.30,3.85,4.13,4.59,4.11,4.29,5.82,5.14)

MLUScores <- data.frame(Sarah,Mother)
```

Calculating the covariance:


```{r}
MLUScores$covariance <- sum((Sarah-mean(Sarah))*(Mother-mean(Mother))/(nrow(MLUScores)-1))

```

Can also be done with a function:

```{r}
cov(Sarah,Mother)
```

Afterwards i can calculate the corralation coefficient

```{r}
MLUScores$CorrelationCoefficient <- (MLUScores$covariance)/(sd(Sarah)*sd(Mother))
```

Visual representation of this 

```{r}
ggplot(MLUScores, aes(Sarah,Mother)) + geom_point() + geom_smooth(method = "lm")
```

The standard correlation as a function:

```{r}
cor.test(Sarah,Mother)
```

Normally distributed data 

The p-value :

Shapiro wilk test: is data different from normal distribution?

p<0.05, yes it is different.

Pearsons test:

if:

p<0.05 yes it has correlation 


5% propability if the null hypothesis is true


Non-normal distributed data:

Spearman:

Ranks data and then applies pearsons equation

Kendalls tau:

Better for smaller samples


```{r}
cor.test(Sarah,Mother)
```


