---
title: "Categorical predictors and multiple regression"
author: "Lasse Hansen"
date: "11/4/2019"
output: html_document
---


Often in regression analysis you’ll collect data about groups of people (e.g., ethnic group, gender, socio-economic status, diagnostic category). You might want to include these groups as predictors in the regression model; however, we saw from our assumptions that variables need to be continuous or categorical with only two categories. 

The obvious problem in wanting to use categorical variables as predictors, is that often you'll have more than two categories.

Here you can use dummy coding.

###Dummy coding

This being R, there are several ways to code dummy variables. We’re going to have a look at the contrasts() function, because we will use it time and time again later in the book. First let’s load the data by executing:

```{r}
gfr <- read.delim("GlastonburyFestivalRegression.dat")
```

Note that the variable music contains text; therefore, R has intelligently decided to cre- ate this variable as a factor, and treat the levels in alphabetical order (level 1 = crusty, level 2 = indie kid, 3 = metaller, and 4 = no musical affiliation). We can use the contrast() function on this variable to set contrasts because it is a factor. There are several built-in contrasts that we can set, we want to compare all groups to a baseline and therefore we execute:

```{r}
contrasts(gfr$music)<-contr.treatment(4, base = 4)
```

The contrasts(gfr$music) simply sets the contrast for the variable music in the gfr dataframe. The contr.treatment() function sets a contrast based on comparing all groups to a baseline.

To set the contrasts manually, because you want to know what contrast 1 is, we create variables that reflect each of the dummy variables:

```{r}
crusty_v_NMA<-c(1, 0, 0, 0) 
indie_v_NMA<-c(0, 1, 0, 0) 
metal_v_NMA<-c(0, 0, 1, 0)
```

Then we need to bind them to the dummy frame:

```{r}
contrasts(gfr$music)<-cbind(crusty_v_NMA, indie_v_NMA, metal_v_NMA)
```

###Regression with dummy variables

Now you’ve created the dummy variables, you can run the regression in the same way as
for any other kind of regression, by executing:

```{r}
Model <- lm(change ~ music, data = gfr)
summary(Model)
```

The fact that we have set contrasts for the variable music means that the three dummy variables will be entered into the model in a single step. This is one reason why setting contrasts is a useful way to handle categorical predictors


ACTUALLY YOU DO NOT HAVE TO DO THE DUMMY CODING BY YOURSELF.


###Interpreting the data

R^2 (multiple r-squared) = This shows that by entering the three dummy variables we can explain 7.6% of the variance in the change in hygiene scores

F-statistic = tells us that the model is significantly better at predicting the change in hygiene scores than having no model (or, put another way, the 7.6% of variance that can be explained is a significant amount)

The beta value tells us the change in the outcome due to a unit change in the predictor. In this case, a unit change in the predictor is the change from 0 to 1. As such it shows the shift in the change in hygiene scores that results from the dummy variable changing from 0 to 1. By including all three dummy variables at the same time, our baseline category is always zero, so this actually represents the difference in the change in hygiene scores if a person has no musical affiliation, compared to someone who is a crusty.