---
title: "Repeated measures design"
author: "Lasse Hansen"
date: "9/25/2019"
output: html_document
---

```{r}
hiccupsData <- read.delim("Hiccups.dat",  header = TRUE)
```

Note there are four variables:
M Baseline: specifies the number of hiccups at baseline.
M Tongue: specifies the number of hiccups after tongue pulling.
M Carotid: specifies the number of hiccups after carotid artery massage. M Rectum: specifies the number of hiccups after digital rectal massage.

Each row in the data file represents a different person, so these data are laid out as a repeated-measures design, with each column representing a different treatment condition and every person undergoing each treatment.
These data are in the wrong format for ggplot2 to use. We need all of the scores stacked up in a single column and then another variable that specifies the type of intervention.

###Rearranging the data

We can rearrange the data as follows (see section 3.9.4):

```{r}
hiccups<-stack(hiccupsData)
names(hiccups)<-c("Hiccups","Intervention")
```

This stacks the data frame into only two rows 

The names() function just assigns names to these new variables in the order that they appear in the dataframe. To plot a categorical variable in ggplot() it needs to be recog- nized as a factor, so we also need to create new variable in the hiccups dataframe called Intervention_Factor, which is just the Intervention variable converted into a factor:

```{r}
hiccups$Intervention_Factor <- factor(hiccups$Intervention)
```

We are now ready to plot the graph. As always we first create the plot object and define the variables that we want to plot as aesthetics:

```{r}
line <- ggplot(hiccups, aes(Intervention_Factor, Hiccups))
```

I have called the object line, and have created it using the ggplot() function. The function specifies the dataframe to be used (hiccups) and has set Intervention_Factor to be plotted on the x-axis, and Hiccups to be plotted on the y-axis.

###Creating the mean value within each treatment

Just as we did for our bar charts, we are going to use stat_summary() to create the mean values within each treatment condition. Therefore, as with the bar chart, we create a layer using stat_summary() and add this to the plot:

```{r}
line + stat_summary(fun.y = mean, geom = "point")
```

Note that this command is exactly the same as for a bar chart, except that we have chosen the point geom rather than a bar. At the moment we have a plot with a symbol representing each group mean. If we want to connect these symbols with a line then we use stat_summary() again, we again specify fun.y to be the mean, but this time choose the line geom. To make the line display we also need to set an aesthetic of group = 1; this is because we are joining summary points (i.e., points that summarize a group) rather than individual data points. Therefore, we specify the line as:

```{r}
line + stat_summary(fun.y = mean, geom = "line", aes(group = 1))
```

Now let’s add an error bar to each group mean. We can do this by adding another layer using stat_summary(). When we plotted an error bar on the bar chart we used a normal error bar, so this time let’s add an error bar based on bootstrapping. We set the function for the data to be mean_cl_boot (fun.data = mean_cl_boot) – see Table 4.4 – and set the geom to be errorbar (you could use pointrange as we did for the bar chart if you prefer):


```{r}
line+ stat_summary(fun.data = mean_cl_boot, geom = "errorbar")
```

###Changing looks of it and adding labels

The default error bars are quite wide, so I recommend setting the width parameter to 0.2 to make them look nicer:

```{r}
line+ stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2)
```

You can, of course, also change the colour and other properties of the error bar in the usual way (e.g., by adding colour = “Red” to make them red). Finally, we will add some labels to the x- and y-axes using the labs() function


```{r}
line+ labs(x = "Intervention", y = "Mean Number of Hiccups")
```

###Full command all together

If we put all of these commands together, we can create the graph by executing the following command:

```{r}
line <- ggplot(hiccups, aes(Intervention_Factor, Hiccups))
line + stat_summary(fun.y = mean, geom = "point") + stat_summary(fun.y = mean, geom = "line", aes(group = 1),colour = "Blue", linetype = "dashed") + stat_summary(fun.data = mean_cl_boot, geom = "errorbar", width = 0.2) + labs(x = "Intervention", y = "Mean Number of Hiccups")
```


###Conclusion of results

The resulting graph in Figure 4.28 displays the mean number of hiccups at baseline and after the three interventions (and the confidence intervals of those means based on boot- strapping). As we will see in Chapter 9, the error bars on graphs of repeated-measures designs aren’t corrected for the fact that the data points are dependent; I don’t want to get into the reasons why here because I want to keep things simple, but if you’re doing a graph of your own data then I would read section 9.2 before you do.
We can conclude that the amount of hiccups after tongue pulling was about the same as at baseline; however, carotid artery massage reduced hiccups, but not by as much as a good old fashioned digital rectal massage. The moral here is: if you have hiccups, find something digital and go amuse yourself for a few minutes.


