---
title: "Regression analysis with single independent variables"
author: "Lasse Hansen"
date: "11/3/2019"
output: html_document
---

###Running a linear regression

```{r}
pacman::p_load(QuantPsyc, boot)
```

Loading data

```{r}
album1 <- read.delim("Album Sales 1.dat")
```

We run a regression analysis with the function lm(), (linear model). It takes the general form:

```{r}
modelName <- lm(outcome ~ predictor, data = dataFrame, na.action = an action))
```

outcome = the variable you are trying the predict (dependent)

Predictor = the variable/variables you are trying to predict from

na.action = If you have missing values it can be used. You can use na.omit or na.exclude, then it will
exclude any case that has missing values.

you do not need dataframe is you use $ like this:

```{r}
albumSales.1 <- lm(album1$sales ~ album1$adverts)
```

###Interpreting a simple regression

We have created an object called albumSales.1 that contains the results of our analysis. We can
show the object by executing:

```{r}
summary(albumSales.1)
```


Let’s start at the bottom of the output:
Multiple R-squared: 0.3346 = R^2

Adjusted R-squared: 0.3313 = Adjusted R^2 for the model that has been derived

We can ofc get R by typing:

```{r}
sqrt(0.3346)
```

The value of R2 of .335 also tells us that advertising expenditure can account for 33.5% of the variation in album sales.

This means that 67% of the variation in album sales can- not be explained by advertising alone. Therefore, there must be other variables that have an influence also.

The next part of the output reports the results of an analysis of variance (ANOVA):
F-statistic: 99.59 on 1 and 198 DF, p-value: < 2.2e-16

It doesn’t give us all of the sums of squares, it just gives the important part: the F-ratio,, and the associated significance value of that F-ratio. For these data, F is 99.59, which is significant at p < .001^4 (because the value labelled p-value is less than .001). This result tells us that there is less than a 0.1% chance that an F-ratio this large would happen if the null hypothesis were true. 

##Coefficients

Intercept = b0, and is 134.1, and this can be interpreted as meaning that
when no money is spent on advertising (when X = 0), the model predicts
that 134,100 albums will be sold (remember that our unit of measurement was thousands of albums)

We can also read off the value of b1 from the row labelled adverts and this value represents the gradient of the regression line. It is 0.096.

Our units of measurement were thousands of pounds and thousands of albums sold, so we can say that for an increase in advertising of £1000 the model predicts 96 (0.096 × 1000 = 96) extra album sales.

###How is the t in the output calculated? Use the values in the output to see if you can get the same value as R.

You calculate it by taking the individual intercept value and dividing it with the standard error.

So T for the value of B1:

```{r}
1.341*10^2/7.537
```

And for B0:

```{r}
(9.612*10^-2)/(9.632*10^-3)
```


###How many units would be sold if we spent £666,000 on advertising the latest album by black metal band Abgott?

We take the b-values in the equation with the values from the output. 

Thereby we can make predictions about album sales, by replacing the advertising budget with a value of interest. 

Here we will use the value 666.000, remembering that out units are already in thousands of pounds we can replace the budget with 100:

```{r}
album.Sales <- 134.14+(0.096*666)
```

This means that album sales whould be around 198 for the first week of sales if they spent 666.000 on
advertising.




