---
title: "Testing if a distribution is normal"
author: "Lasse Hansen"
date: "9/24/2019"
output: html_document
---
####The Shapiro wilk test

If the test is non-significant (p > .05) it tells us that the distribution of the sample is not significantly different from a normal distribution. If, however, the test is significant (p < .05) then the distribution in question is significantly differ- ent from a normal distribution (i.e., it is non-normal).

###Limitations of the test 

However, it has limitations because with large sample sizes it is very easy to get significant results from small deviations from normality, and so a
 significant test doesn’t necessarily tell us whether the deviation from normality is enough to bias any statistical procedures that we apply to the data. I guess the take-home message is: by all means use these tests, but plot your data as well and try to make an informed decision about the extent of non-normality.
 
This function takes the general form: 

```{r}
shapiro.test(variable)
```
in which variable is the name of the variable that you’d like to test for normality. Therefore,
to test the exam and numeracy variables for normality we would execute:

```{r}
shapiro.test(rexam$exam)
shapiro.test(rexam$numeracy)
```

Remember that a significant value (p-value less than .05) indicates a deviation from normality. For both numeracy (p = .005) and R exam scores (p < .001), the Shapiro–Wilk test is highly significant, indicating that both distributions are not normal. 

Now if we’d asked for separate Shapiro– Wilk tests for the two universities we might have found non-significant results. 

We use the by() function again

```{r}
by(rexam$exam, rexam$uni, shapiro.test)
by(rexam$numeracy, rexam$uni, shapiro.test)
```

This is important because if our analysis involves comparing groups, then what’s important is not the overall distribution but the distribution in each group.


###The test statistic for the Shapiro–Wilk test is denoted by W; we can report the results in. Output 5.5 in the following way:

The percentage on the R exam, W = 0.96, p = .005, and the numeracy scores, W =
0.92, p < .001, were both significantly non-normal.


