---
title: "The dependent t-test"
author: "Lasse Hansen"
date: "10/30/2019"
output: html_document
---

```{r}
install.packages(c("MASS", "akima", "robustbase"))
install.packages(c("cobs", "robust", "mgcv", "scatterplot3d", "quantreg", "rrcov", "lars", "pwr", "trimcluster", "parallel", "mc2d", "psych", "Rfit"))
install.packages("WRS", repos="http://R-Forge.R-project.org", type="source")

```


```{r}
pacman::p_load(tidyverse, ggplot, pastecs, WRS)

```

```{r}
picture<-c(30, 35, 45, 40, 50, 35, 55, 25, 30, 45, 40, 50)
real<-c(40, 35, 50, 55, 65, 55, 50, 35, 30, 50, 60, 39)

Spider.Data <- data.frame(picture,real)

```

Checking for normality

```{r}
stat.desc(Spider.Data, basic = FALSE, norm = TRUE)
```

Yes the data looks normal, it would be optimal to visualise it with a ggplot.

In this case we have scores from different groups stored in different
columns, so we could execute paired = TRUE:

```{r}
Dependent.Ttest <- t.test(Spider.Data$real, Spider.Data$picture, paired =  TRUE)
Dependent.Ttest
```

##Explaining  the results

The size of t is compared against known values based on the degrees of freedom. When the same participants have been used, the degrees of freedom are simply the sample size minus 1 (df = N −1 = 11) – you should check this value is what you expect it to be, to ensure you haven’t made a mistake. R uses the degrees of freedom to calculate the exact probability that a value of t as big as the one obtained could occur if the null hypothesis were true (i.e., there was no difference between these means).

The probability for the spider data is very low (p = .031) and in fact it tells us that there is only a 3.1% chance that a value of t this big could happen if the null hypothesis were true. We generally accept a 
p < .05 as statistically meaningful; therefore, this t is significant because .031 is smaller than .05.

The fact that the t-value is a positive number tells us that the first condition (the real condition) had a larger mean than the second (the picture condition) and so the real spider led to greater anxiety than the picture. Therefore, we can conclude that exposure to a real spider caused significantly more reported anxiety in arachnophobes than exposure to a picture, t(11) = 2.47, p < .05.


###Robust measures to compare dependent means

The first robust function, yuend(), is based on a trimmed mean. It takes the general form:

yuend(scores group 1, scores group 2, tr = .2, alpha = .05)

As such, for a test of dependent means based on 20% trimming we simply execute:

```{r}
yuend(Spider.Data$real,Spider.Data$picture, tr = .2, alpha = .05)
```

Based on this robust test there is not a significant difference in anxiety scores across the two spider groups, Ty(7) = 1.86, p = .106.


###Calculating the effect size

Even though our t-statistic is statistically significant, this doesn’t mean our effect is important in practical terms. To find out if the effect is substansive we use our knowledge on effect sizes:

```{r}
t <- Dependent.Ttest$statistic[[1]]
df <- Dependent.Ttest$parameter[[1]]
r <- sqrt(t^2/(t^2+df))
round(r,  3)
```

This represents a very large effect according to the threshold for a large effect. Therefore, as well as being statistically significant, this effect is large and probably substantive finding.

###Reporting  the dependent t-test

In this example the R output tells us that the value of t was 2.47, that this was based on 11 degrees of freedom, and that it was significant at p = .031. We can also see the means for each group. We could write this as:

On average, participants experienced significantly greater anxiety from real spiders (M = 47.00, SE = 3.18) than from pictures of spiders (M = 40.00, SE = 2.68), t(11) = 2.47, p < .05, r = .60.


