---
title: "When variances are not equal across groups"
author: "Lasse Hansen"
date: "11/14/2019"
output: html_document
---


Packages:

```{r}
pacman::p_load(compute.es,car,ggplot2,multcomp,pastecs,WRS)
```

To conduct ANOVA, follow this procedure:

1. Enter data
2. Explore your data (checking assumptions)
3. Compute the basic ANOVA (run the main analysis, robust if needed)
4. Compute contrasts or post hoc tests

###Entering data

The data must be entered in two columns. We are using data on viagra here.
The grouping variables are coded so that 1 = placebo, 2 = low dose, 3 = high dose:

```{r}
libido<-c(3,2,1,1,4,5,2,4,2,3,7,4,5,3,6)
dose<-gl(3,5, labels = c("Placebo", "Low Dose", "High Dose"))
viagraData<-data.frame(dose, libido)
```

gl() makes  the dose column a factor variable.

###Levenes test

If Levene’s test is significant then it is reasonable to assume that population variances are different across groups.

###Welch's F

Welch’s F, makes adjustments for differences in group variances. This test is produced by the oneway.test() function, which is built into R. The format of this test is the same as aov():

```{r}
oneway.test(viagraData$libido ~ viagraData$dose)
```

Welch’s F(2, 7.94) = 4.23, p = .054, which is just about non-significant. If we were using this test it would imply that the mean libido did not differ significantly across different doses of Viagra.

